---
title: "Calculations- working"
author: "Katie Jolly"
date: "February 28, 2018"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: "cosmo"
---

```{r include = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```


# Project setup

```{r}
# packages

library(raster)
library(rgdal)
library(sp)
library(leaflet)
library(tigris)
library(tidycensus)
library(tidyverse)
library(janitor)
library(gridExtra)
library(ggthemes)
library(spdep)

# cache the tigris API calls
options(tigris_use_cache = TRUE)
```

## Spatial data: loading + API calls

The tigris data is cached, but not the shapefile. Think about caching all in the future.

```{r}
ca_schools_shp <- readOGR("C:\\Users\\katie\\Documents\\healthGIS_spatial_lag_modeling\\california_schools_pbe\\CA_schools_PBE.shp")

ca_school_dist_tigris <- school_districts(state = "CA")
ca_tract_tigris <- tracts(state = "CA", year = 2015)
ca_block_tigris <- block_groups(state = "CA", year = 2015)
```


## Set CRS

I used California Teale Albers (official state projection) to project the data. `Proj4` found [here](http://spatialreference.org/ref/sr-org/10/proj4/). Guidance found in Nick Eubank's [post]()

```{r}
common_crs <- CRS("+proj=aea +lat_1=34 +lat_2=40.5 +lat_0=0 +lon_0=-120 +x_0=0 +y_0=-4000000 +ellps=GRS80 +datum=NAD83 +units=m +no_defs +init=epsg:26911")

ca_tract_tigris_proj <- spTransform(ca_tract_tigris, common_crs)
ca_block_tigris_proj <- spTransform(ca_block_tigris, common_crs)
ca_school_dist_tigris_proj <- spTransform(ca_school_dist_tigris, common_crs)

ca_schools_shp_proj <- spTransform(ca_schools_shp, common_crs)
```

# Spatial joins

Did a spatial join with the school location data and each of the three levels of geography. 

```{r}
ca_dist_join <- over(ca_schools_shp_proj, ca_school_dist_tigris_proj) # schools + districts
ca_schools_shp_proj@data <- data.frame(ca_schools_shp_proj@data, ca_dist_join) # set the new data


ca_tract_join <- over(ca_schools_shp_proj, ca_tract_tigris_proj) # schools + tracts
ca_schools_shp_proj@data <- data.frame(ca_schools_shp_proj@data, ca_tract_join) # set the new data

ca_block_join <- over(ca_schools_shp_proj, ca_block_tigris_proj) # schools + block
ca_schools_shp_proj@data <- data.frame(ca_schools_shp_proj@data, ca_block_join) # set the new data

schools_data_full <- tibble() # create an empty tibble for the data because otherwise it doesn't clear when you run it again

schools_data_full <- ca_schools_shp_proj@data %>% # pull out the data to work with 
  as_tibble()
```

# Descriptive statistics

```{r}
schools_data <- schools_data_full %>% # take only the important variables
  dplyr::select(c(CDSCode, Street, City, Zip, County_1, PublicPriv, Name, LatDec, LonDec, EnrollTot, PBETot, GEOID, NAME, LOGRADE, HIGRADE, GEOID.1, GEOID.2)) %>% # rename some of the ambigous variables
  rename(county = County_1,
         latitute = LatDec,
         longitude = LonDec,
         geoid_dist = GEOID, 
         dist_name = NAME,
         lo_grade = LOGRADE,
         hi_grade = HIGRADE,
         geoid_tract = GEOID.1,
         geoid_block = GEOID.2) %>%
  clean_names() # make them all lowercase for consistency

nlevels(factor(schools_data$geoid_dist)) # number of school districts
nlevels(factor(schools_data$geoid_tract)) # number of census tracts
nlevels(factor(schools_data$geoid_block)) # number of block groups
```

```{r}
geo_summarise <- function(group_by) {
  group_by <- enquo(group_by) # make this useful as a variable to have in dplyr functions

  schools_data %>%
    group_by(!!group_by) %>% # unquote the group_by variable
    summarise(pbe = sum(pbetot, na.rm = TRUE), # total number of exemptions
            enroll = sum(enrolltot, na.rm = TRUE), # total enrollment
            pbe_rate = pbe/enroll) # geo-wide pbe rate
}

by_district <- geo_summarise(geoid_dist) # using school district as the geography
by_tract <- geo_summarise(geoid_tract) # using census tracts as the geography
by_block <- geo_summarise(geoid_block) # using block groups as the geography
```

```{r}
blue <- "#6b8ec6"
yellow <- "#e4e888"

theme <- theme(axis.title = element_text())

dist_density <- ggplot(by_district, aes(x = pbe_rate)) +
  geom_density(fill = blue, color = "gray10") +
  theme_hc() +
  theme +
  labs(y = NULL,
       x = "PBE rate, school district level")

tract_density <- ggplot(by_tract, aes(x = pbe_rate)) +
  geom_density(fill = blue, color = "gray10") +
  theme_hc() +
  theme +
  labs(y = NULL,
       x = "PBE rate, census tract level")

block_density <- ggplot(by_block, aes(x = pbe_rate)) +
  geom_density(fill = blue, color = "gray10") +
  theme_hc() +
  theme +
  labs(y = NULL,
       x = "PBE rate, block group level")

grid.arrange(dist_density, tract_density, block_density, ncol = 1, nrow = 3)
```

Observations:

* very right skewed
  * there are a few geographies with relatively right rates of exemptions
  * in this case, outliers are very important!
* general shape is the same, but school district rates are less concentrated
  * likely because the exemptions are spread out over higher enrollment numbers
  
```{r}
# district_ogive <- ggplot(by_district, aes(x = pbe_rate)) +
#   stat_ecdf(color = blue, size = 1.7) +
#   theme_fivethirtyeight() +
#   theme +
#   labs(y = NULL,
#        x = "Cumulative PBE frequency, school district level")
# 
# tract_ogive <- ggplot(by_tract, aes(x = pbe_rate)) +
#   stat_ecdf(color = blue, size = 1.7) +
#   theme_fivethirtyeight() +
#   theme +
#   labs(y = NULL,
#        x = "Cumulative PBE frequency, census tract level")
# 
# block_ogive <- ggplot(by_block, aes(x = pbe_rate)) +
#   stat_ecdf(color = blue, size = 1.7) +
#   theme_fivethirtyeight() +
#   theme +
#   labs(y = NULL,
#        x = "Cumulative PBE frequency, block group level")

```

Add more descriptive statistics later!

# Calculate Monte Carlo Moran's I

```{r}
ca_tract_tigris_join <- ca_tract_tigris %>%
  geo_join(by_tract, by_sp = "GEOID", by_df = "geoid_tract")


ca_district_tigris_join <- ca_school_dist_tigris %>%
  geo_join(by_district, by_sp = "GEOID", by_df = "geoid_dist")

leaflet(ca_schools_shp) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addMarkers(
  clusterOptions = markerClusterOptions()
)
```

```{r}
ca_schools_shp$pbe_rate <- ca_schools_shp$PBETot/ca_schools_shp$EnrollTot # calculate the pbe rates per school

schools_knn <- knearneigh(ca_schools_shp, k = 10) # neighborhood size 5

schools_nb <- knn2nb(schools_knn) # convert to nb object from knn

swm <- nb2mat(schools_nb, style='B') # this makes a spatial weights matrix-- useful but not what I'm looking for

ww <-  nb2listw(schools_nb, style='B') # makes another weights matrix, not sure how it's different but apparently it's what I'm looking for

mc_points <- moran.mc(ca_schools_shp$pbe_rate, ww, nsim = 99) # simulate 99 + 1 times

mc_points # summary of the results

mc_values <- as_tibble(x = mc_points$res) # results of each simulation

ggplot(mc_values, aes(x = value)) +
  geom_density(fill = "gray50") +
  geom_vline(xintercept = mc_points$statistic, color = blue, size = 1.5) +
  theme_hc() +
  labs(y = NULL, x = "Moran's I index values",
       title = "Distribution of expected Moran's I statistics using k-nearest neighbors",
       subtitle = "Neighborhood size (k) = 5") +
  annotate("text", x = 0.23, y = 50, label = "Computed statistic")

moran_knn_schools <- function(k = 10, nsim = 99){
  knn <- knearneigh(ca_schools_shp, k = k)
  nb <- knn2nb(knn) 
  ww <- nb2listw(nb, style = 'B')
  moran <- moran.mc(ca_schools_shp$pbe_rate, ww, nsim = nsim)
}

stats <- c()
p <- c()

for (k in 1:100){
  knn <- moran_knn_schools(k = k)
  stats <- c(stats, knn$statistic)
  p <- c(p, knn$p.value)
}

stats # vector of statistics (index values)
p # vector of p-values

moran_results <- tibble(index = 1:100,
                        value = stats,
                        p = p) # make it all into a tibble with index values

ggplot(moran_results, aes(x = index, y = value)) +
  geom_point(color = blue) +
  labs(y = "Moran's I statistic",
       x = "k (neighborhood size)",
       title = "Spatial Autocorrelation of Personal Belief Exemptions",
       subtitle = "Results of computing Monte Carlo Moran's I with k neighbors, 6469 California kindergartens") +
  theme_hc()

```

# Data coverage

For which school districts do we have data?

Join problem with the data!!! 

```{r}
crs_leaflet <- leafletCRS(proj4def = "+proj=aea +lat_1=34 +lat_2=40.5 +lat_0=0 +lon_0=-120 +x_0=0 +y_0=-4000000 +ellps=GRS80 +datum=NAD83 +units=m +no_defs +towgs84=0,0,0 +init=epsg:26911")

leaflet(ca_district_tigris_join) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(fillColor = blue,
  weight = 0,
  opacity = 0,
  color = NULL,
  dashArray = NULL,
  fillOpacity = 0.9) %>%
  addMarkers(data = ca_schools_shp,
  clusterOptions = markerClusterOptions(),
  label = ~as.character(Name)
)
```

For which census tracts do we have data?

```{r}
ca_tract_tigris_join <- subset(ca_tract_tigris_join, !is.na(pbe_rate)) # take out NA census tracts

leaflet(ca_tract_tigris_join) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(fillColor = blue,
  weight = 0,
  opacity = 0,
  color = NULL,
  dashArray = NULL,
  fillOpacity = 0.9)
```

